### 2022-02-19 paper/output/document/thesis_3_implementation.tex
i wrote this chapter in a day lol

%
\begin{isabellebody}%
\setisabellecontext{thesis{\isacharunderscore}{\isadigit{3}}{\isacharunderscore}implementation}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Implementation Details\label{details}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In this section, I present the details of my implementation of automated Kantian ethics, which requires
formalizing the FUL in Dyadic Deontic Logic and then implementing this logic in Isabelle/HOL. The final
Isabelle library contains enough logical background and an axiom representing the categorical imperative. 
Using Isabelle's automated theorem proving abilities, my system can show that appropriately represented
actions are, according to Kantian ethics, obligatory, permissible, or prohibited by proving or refuting sentences of the form ``A 
is obligated to do B.'' I also present a testing framework to evaluate how faithful my implementation is 
to philosophical literature. This testing framework shows that my system outperforms two other potential 
implementations of automated Kantian ethics.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Formalization and Implementation of the FUL%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
Formalizing the FUL requires defining and implementing enough logical background to represent 
the FUL as an axiom. Dyadic
Deontic Logic can express obligation and prohibition, but it cannot represent more complex features of 
moral judgement like actions, subject, maxims, and ends. I augment DDL by adding representations 
of these concepts, drawn from philosophical literature.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{Logical Background%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
Kantian ethics is fundamentally action-guiding; the categorical
imperative is a moral rule that agents can use to decide between potential actions. Thus, before I 
begin to formalize a specific formulation of the categorical imperative, I must define the notions of 
subjects and actions. I add representations of subjects and actions so that my new logic can express
sentences of the form, ``x does action.''%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{typedecl}\isamarkupfalse%
\ s\ %
\isamarkupcmt{The new type $s$ is the type for a ``subject,'' as in the subject of a sentence.%
}%
\begin{isamarkuptext}%
The \texttt{typedecl} keyword indicates that I am defining a new atomic type, which is not composed
of pre-existing types but is instead a new kind of object altogether. It suffices
to declare a new type to represent a subject without specifying any of its complex properties, such as the idea that a subject
must be rational or human. Instead of providing a complete definition of subject, which would require wading into murky
philosophical debates about the nature of agency, I will provide a ``thin'' definition that only includes
the minimum necessary properties to apply the FUL. Throughout my project, I will use bare syntactic units like types and constants
to create thin definitions of new ideas.

@@

@@

%:%113=47%:%
%:%115=48%:%
%:%116=49%:%
%:%117=50%:%
%:%118=51%:%
%:%119=52%:%

@@

%:%130=57%:%
%:%142=59%:%
%:%143=60%:%
%:%144=61%:%
%:%145=62%:%
%:%146=63%:%

@@

%:%170=88%:%

@@

%:%173=89%:%
%:%174=90%:%
%:%175=91%:%
%:%176=92%:%
%:%177=93%:%

@@

%:%183=98%:%
%:%184=99%:%
%:%185=100%:%
%:%186=101%:%
%:%187=102%:%

@@

@@

%:%243=157%:%
%:%244=158%:%
%:%246=159%:%
%:%247=160%:%
%:%248=161%:%
%:%249=162%:%
%:%257=164%:%
%:%269=166%:%
%:%270=167%:%
%:%271=168%:%
%:%272=169%:%
%:%273=170%:%

@@

%:%321=219%:%

@@

%:%325=221%:%
%:%326=222%:%
%:%329=224%:%
%:%330=225%:%
%:%331=226%:%
%:%332=227%:%
%:%333=228%:%

@@

%:%352=248%:%

@@

%:%356=250%:%
%:%357=251%:%
%:%358=251%:%
%:%361=257%:%
%:%362=258%:%
%:%363=259%:%
%:%364=260%:%
%:%365=261%:%

@@

@@

%:%374=269%:%
%:%375=270%:%
%:%377=271%:%
%:%380=273%:%
%:%381=274%:%
%:%382=275%:%
%:%383=276%:%
%:%384=277%:%

@@

@@

%:%415=307%:%
%:%416=308%:%
%:%418=309%:%
%:%419=310%:%
%:%420=310%:%
%:%421=311%:%
%:%422=312%:%
%:%423=312%:%
%:%424=313%:%
%:%426=314%:%
%:%427=315%:%
%:%430=317%:%
%:%431=318%:%
%:%432=319%:%
%:%433=320%:%

@@

%:%436=322%:%
%:%437=323%:%
%:%439=324%:%
%:%440=325%:%
%:%441=325%:%
%:%442=326%:%
%:%443=327%:%
%:%444=327%:%
%:%446=328%:%
%:%447=329%:%
%:%448=330%:%
%:%456=332%:%
%:%468=334%:%
%:%469=335%:%
%:%470=336%:%
%:%471=337%:%
%:%472=338%:%

@@

%:%477=344%:%

@@

%:%480=345%:%
%:%481=346%:%
%:%482=347%:%
%:%483=348%:%
%:%486=350%:%
%:%487=351%:%
%:%489=353%:%
%:%490=353%:%
%:%492=353%:%
%:%496=353%:%
%:%497=353%:%
%:%498=354%:%
%:%499=354%:%
%:%500=354%:%
%:%509=356%:%
%:%510=357%:%
%:%511=358%:%
%:%512=359%:%
%:%513=360%:%

@@

@@

%:%536=382%:%
%:%537=383%:%
%:%539=384%:%
%:%540=385%:%
%:%543=387%:%
%:%545=389%:%
%:%546=389%:%
%:%547=390%:%
%:%549=391%:%
%:%550=392%:%
%:%551=392%:%
%:%552=393%:%
%:%553=394%:%
%:%554=394%:%
%:%555=395%:%
%:%556=395%:%
%:%558=395%:%
%:%562=395%:%
%:%563=395%:%
%:%565=396%:%
%:%566=397%:%
%:%567=398%:%
%:%568=399%:%
%:%569=400%:%

@@

%:%581=404%:%
%:%582=405%:%
%:%583=406%:%
%:%584=407%:%
%:%585=408%:%

@@

@@

%:%604=426%:%
%:%606=428%:%
%:%607=429%:%
%:%608=430%:%

@@

%:%610=432%:%
%:%619=434%:%
%:%631=436%:%
%:%632=437%:%
%:%633=438%:%
%:%634=439%:%
%:%635=440%:%

@@

%:%658=464%:%
%:%659=465%:%
%:%660=466%:%
%:%661=467%:%
%:%662=468%:%

@@

%:%666=471%:%
%:%667=472%:%
%:%668=473%:%
%:%669=474%:%
%:%670=475%:%

@@

%:%679=482%:%
%:%680=483%:%
%:%681=484%:%
%:%682=485%:%
%:%683=486%:%

@@

@@

%:%695=497%:%
%:%696=498%:%
%:%699=499%:%
%:%703=499%:%
%:%704=499%:%
%:%706=500%:%
%:%707=501%:%
%:%708=502%:%
%:%709=502%:%
%:%711=503%:%
%:%712=504%:%
%:%713=505%:%
%:%714=506%:%
%:%720=506%:%
%:%723=507%:%
%:%724=508%:%
%:%725=508%:%
%:%726=509%:%
%:%729=510%:%
%:%733=510%:%
%:%734=510%:%
%:%736=511%:%
%:%737=512%:%
%:%747=514%:%
%:%748=515%:%
%:%749=516%:%
%:%750=517%:%
%:%751=518%:%

@@

@@

%:%770=536%:%
%:%771=537%:%
%:%774=538%:%
%:%778=538%:%
%:%779=538%:%
%:%780=538%:%
%:%782=539%:%
%:%788=539%:%
%:%791=540%:%
%:%792=541%:%
%:%793=541%:%
%:%794=542%:%
%:%795=543%:%
%:%798=544%:%
%:%802=544%:%
%:%803=544%:%
%:%804=544%:%
%:%806=545%:%
%:%807=546%:%
%:%808=547%:%
%:%818=549%:%
%:%819=550%:%
%:%820=551%:%
%:%821=552%:%
%:%822=553%:%

@@

%:%827=559%:%

@@

%:%831=560%:%
%:%835=560%:%
%:%836=560%:%
%:%838=561%:%
%:%839=562%:%
%:%840=563%:%
%:%850=565%:%
%:%851=566%:%
%:%852=567%:%
%:%853=568%:%
%:%854=569%:%

@@

@@

%:%859=573%:%
%:%860=574%:%
%:%861=575%:%
%:%864=576%:%
%:%868=576%:%
%:%869=576%:%
%:%870=576%:%
%:%872=577%:%
%:%873=578%:%
%:%874=579%:%
%:%884=581%:%
%:%885=582%:%
%:%887=584%:%
%:%888=584%:%
%:%889=585%:%
%:%890=586%:%
%:%893=587%:%
%:%897=587%:%
%:%898=587%:%
%:%899=587%:%
%:%901=588%:%
%:%911=590%:%
%:%912=591%:%
%:%913=592%:%
%:%914=593%:%
%:%915=594%:%

@@

@@

%:%930=609%:%
%:%931=610%:%
%:%932=611%:%

@@

%:%939=617%:%
%:%940=618%:%
%:%941=619%:%
%:%942=620%:%
%:%943=621%:%

@@

>>>>

\begin{isabellebody}%
\setisabellecontext{thesis{\isacharunderscore}{\isadigit{3}}{\isacharunderscore}implementation}%
\isamarkupsection{Implementation Details\label{details}%
In this section, I present the details of my implementation of automated Kantian ethics, which requires
formalizing the FUL in Dyadic Deontic Logic and then implementing this logic in Isabelle/HOL. The final
Isabelle library contains enough logical background and an axiom representing the categorical imperative. 
Using Isabelle's automated theorem proving abilities, my system can show that appropriately represented
actions are, according to Kantian ethics, obligatory, permissible, or prohibited by proving or refuting sentences of the form ``A 
is obligated to do B.'' I also present a testing framework to evaluate how faithful my implementation is 
to philosophical literature. This testing framework shows that my system outperforms two other potential 
implementations of automated Kantian ethics.%
\isamarkupsubsection{Formalization and Implementation of the FUL%
Formalizing the FUL requires defining and implementing enough logical background to represent 
the FUL as an axiom. Dyadic
Deontic Logic can express obligation and prohibition, but it cannot represent more complex features of 
moral judgement like actions, subject, maxims, and ends. I augment DDL by adding representations 
of these concepts, drawn from philosophical literature.%
\isamarkupsubsubsection{Logical Background%
Kantian ethics is fundamentally action-guiding; the categorical
imperative is a moral rule that agents can use to decide between potential actions. Thus, before I 
begin to formalize a specific formulation of the categorical imperative, I must define the notions of 
subjects and actions. I add representations of subjects and actions so that my new logic can express
sentences of the form, ``x does action.''%
\isacommand{typedecl}\isamarkupfalse%
\ s\ %
\isamarkupcmt{The new type $s$ is the type for a ``subject,'' as in the subject of a sentence.%
The \texttt{typedecl} keyword indicates that I am defining a new atomic type, which is not composed
of pre-existing types but is instead a new kind of object altogether. A type does not come with any 
properties out of the box. There is no difference between declaring the type ``subject'' or any other
type, such as ``color'' or ``mammal.'' I add some properties of this type below by creating formulas
and more complex types that use this type, but I do not provide anything near a definition of a subject.
It suffices to declare a new type to represent a subject without specifying any of its complex properties, 
such as the idea that a subject must be rational or human. Instead of providing a complete definition of subject, 
which would require wading into murky philosophical debates about the nature of agency, I provide a 
``thin'' definition that only includes the minimum necessary properties to apply the FUL. Throughout my 
project, I will use bare syntactic units like types and constants to create thin definitions of new ideas.
This strategy lets me avoid messy philosophical controversies and makes my system's judgements more 
trustworthy, because they rely on relatively little prior knowledge.

@@

%:%111=46%:%

@@

%:%113=48%:%
%:%114=49%:%
%:%115=50%:%
%:%117=52%:%
%:%118=52%:%

@@

%:%123=56%:%
%:%124=57%:%
%:%125=58%:%
%:%126=59%:%
%:%127=60%:%
%:%135=62%:%

@@

%:%169=86%:%
%:%170=87%:%

@@

%:%172=89%:%
%:%173=90%:%
%:%175=93%:%
%:%176=93%:%

@@

%:%181=97%:%
%:%182=98%:%
%:%183=99%:%
%:%184=100%:%
%:%185=101%:%

@@

%:%241=156%:%

@@

%:%243=158%:%
%:%244=159%:%
%:%245=160%:%
%:%247=162%:%
%:%248=162%:%
%:%249=163%:%
%:%251=164%:%
%:%252=165%:%
%:%253=166%:%
%:%254=167%:%
%:%262=169%:%

@@

%:%320=217%:%
%:%321=218%:%

@@

%:%324=221%:%
%:%326=224%:%
%:%327=224%:%
%:%328=225%:%
%:%330=226%:%
%:%331=227%:%

@@

%:%351=246%:%
%:%352=247%:%

@@

%:%355=250%:%
%:%357=253%:%
%:%358=253%:%
%:%359=254%:%
%:%361=255%:%
%:%362=256%:%
%:%363=256%:%

@@

%:%372=268%:%

@@

%:%374=270%:%
%:%375=271%:%
%:%376=272%:%
%:%378=274%:%
%:%379=274%:%
%:%380=275%:%
%:%382=276%:%

@@

%:%413=306%:%

@@

%:%415=308%:%
%:%416=309%:%
%:%417=310%:%
%:%419=312%:%
%:%420=312%:%
%:%421=313%:%
%:%423=314%:%
%:%424=315%:%
%:%425=315%:%
%:%426=316%:%
%:%427=317%:%
%:%428=317%:%
%:%429=318%:%
%:%431=319%:%
%:%432=320%:%

@@

%:%436=323%:%
%:%437=324%:%
%:%438=325%:%
%:%440=327%:%
%:%441=327%:%
%:%442=328%:%
%:%444=329%:%
%:%445=330%:%
%:%446=330%:%
%:%447=331%:%
%:%448=332%:%
%:%449=332%:%
%:%451=333%:%
%:%452=334%:%
%:%453=335%:%
%:%461=337%:%

@@

%:%476=342%:%
%:%477=343%:%

@@

%:%479=345%:%
%:%480=346%:%
%:%482=349%:%
%:%483=349%:%
%:%485=350%:%
%:%486=351%:%
%:%487=352%:%
%:%488=353%:%
%:%491=355%:%
%:%492=356%:%
%:%494=358%:%
%:%495=358%:%
%:%497=358%:%
%:%501=358%:%
%:%502=358%:%
%:%503=359%:%
%:%504=359%:%
%:%505=359%:%

@@

%:%534=381%:%

@@

%:%536=383%:%
%:%537=384%:%
%:%538=385%:%
%:%540=387%:%
%:%541=387%:%
%:%542=388%:%
%:%544=389%:%
%:%545=390%:%
%:%548=392%:%
%:%550=394%:%
%:%551=394%:%
%:%552=395%:%
%:%554=396%:%
%:%555=397%:%
%:%556=397%:%
%:%557=398%:%
%:%558=399%:%
%:%559=399%:%
%:%560=400%:%
%:%561=400%:%
%:%563=400%:%
%:%567=400%:%
%:%568=400%:%

@@

%:%572=403%:%
%:%573=404%:%
%:%574=405%:%
%:%575=406%:%
%:%576=407%:%

@@

%:%602=425%:%

@@

%:%604=427%:%
%:%605=428%:%
%:%606=429%:%
%:%608=431%:%

@@

%:%611=433%:%
%:%612=434%:%
%:%613=435%:%
%:%614=436%:%
%:%615=437%:%
%:%624=439%:%

@@

%:%656=461%:%
%:%657=462%:%
%:%658=463%:%
%:%659=464%:%
%:%660=465%:%

@@

%:%664=470%:%
%:%665=471%:%
%:%666=472%:%
%:%667=473%:%
%:%668=474%:%

@@

%:%676=481%:%
%:%677=482%:%
%:%678=483%:%
%:%679=484%:%
%:%680=485%:%

@@

%:%693=496%:%

@@

%:%695=498%:%
%:%696=499%:%
%:%697=500%:%
%:%699=502%:%
%:%700=502%:%
%:%701=503%:%
%:%704=504%:%
%:%708=504%:%
%:%709=504%:%
%:%711=505%:%
%:%712=506%:%
%:%713=507%:%
%:%714=507%:%
%:%716=508%:%
%:%717=509%:%
%:%718=510%:%
%:%719=511%:%
%:%725=511%:%
%:%728=512%:%
%:%729=513%:%
%:%730=513%:%
%:%731=514%:%
%:%734=515%:%
%:%738=515%:%
%:%739=515%:%
%:%741=516%:%
%:%742=517%:%

@@

%:%768=535%:%

@@

%:%770=537%:%
%:%771=538%:%
%:%772=539%:%
%:%774=541%:%
%:%775=541%:%
%:%776=542%:%
%:%779=543%:%
%:%783=543%:%
%:%784=543%:%
%:%785=543%:%
%:%787=544%:%
%:%793=544%:%
%:%796=545%:%
%:%797=546%:%
%:%798=546%:%
%:%799=547%:%
%:%800=548%:%
%:%803=549%:%
%:%807=549%:%
%:%808=549%:%
%:%809=549%:%
%:%811=550%:%
%:%812=551%:%
%:%813=552%:%

@@

%:%826=557%:%
%:%827=558%:%

@@

%:%829=560%:%
%:%830=561%:%
%:%832=564%:%
%:%833=564%:%
%:%836=565%:%
%:%840=565%:%
%:%841=565%:%
%:%843=566%:%
%:%844=567%:%
%:%845=568%:%

@@

%:%857=572%:%

@@

%:%859=574%:%
%:%860=575%:%
%:%861=576%:%
%:%863=578%:%
%:%864=578%:%
%:%865=579%:%
%:%866=580%:%
%:%869=581%:%
%:%873=581%:%
%:%874=581%:%
%:%875=581%:%
%:%877=582%:%
%:%878=583%:%
%:%879=584%:%
%:%889=586%:%
%:%890=587%:%
%:%892=589%:%
%:%893=589%:%
%:%894=590%:%
%:%895=591%:%
%:%898=592%:%
%:%902=592%:%
%:%903=592%:%
%:%904=592%:%
%:%906=593%:%

@@

%:%925=604%:%
%:%926=605%:%
%:%927=606%:%

@@

@@

%:%936=615%:%
%:%937=616%:%
%:%938=617%:%
%:%939=618%:%
%:%940=619%:%

@@

%:%958=636%:%
%:%959=637%:%
%:%960=638%:%
%:%961=639%:%
%:%962=640%:%
